### 2011-12-14 doc/theses/mikula/pre_creation_of_algorithm.tex
Master thesis: In progress.

Another step can be evaluation of the queries using the XML data and consecutive analysis of the results. And even the process of evaluation itself can be analyzed to obtain some useful information. \todo[inline]{Q: Toto skusit rozpisat nejak podrobnejsie? A: Ty useful information by neškodilo popsat, jinak OK.}

@@

\section{Forms of Query Precessing}
Another important question is how the queries can be processed. Will they be just searched for certain patterns like it is performed in method \cite{Necasky:2009:DXK:1529282.1529414} or will they be processed in a more sophisticated way? That can mean incorporating lexical and syntactic analyses \todo{Q: referencie?}, known from creation of compilers, or even a form of an analysis of semantics \todo{referencia?}.

\todo[inline]{Q: referencie? A: Možná jednu společnou na nějakou základní klinu o překadačích. Toto asi nebude jádro práce, takže není třeba přesně referencovat každý pojem.}

The result of lexical and syntantic analyses can be some kind of so-called syntax tree \todo{referencia}. It is a structure representing a word according to a formal grammar of some language \todo{Q: definovat grammar, tree, language? A: Pokud bude použito dále, tak ano.}. In our case, the language is XQuery, its grammar is defined in \todo{ref} and every query is a word of the XQuery language. Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals. \todo{priklad nejakeho stromu?} From the point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient. For instance, the tree can simplify a search for FLWOR statements. It is transitioned and nodes representing FLWORs are found. Then each subtree determined by one of the found nodes represents a FLWOR statement and it can be analyzed further.

The syntax tree can be also extended by additional information. An example is a static analysis of expression types. Types of literal expressions are defined, functions have return types, path expressions can return nodes, etc. Applying \todo{divná věta} rules defined in TODO \todo{ref} can be determined types of complex expressions. These can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.

@@

A more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions. But it is also more difficult in terms of implementation. On the other hand, simplier approaches of the query processing such as the pattern finding are easier to implement but they also limit possibilities of the query utilization.

@@

\todo[inline]{Pre co sme sa rozhodli? Odvozovat strukturu čistě z dotazů je nesmysl.}

>>>>

Another step can be evaluation of the queries using the XML data and consecutive analysis of the results. And even the process of evaluation itself can be analyzed to obtain some useful information. For instance, these are partial results of evaluating of expressions (elements selected by each path expression, real arguments in fuction calls, etc).

@@

\section{Forms of Query Precessing}
Another important question is how the queries can be processed. Will they be just searched for certain patterns like it is performed in method \cite{Necasky:2009:DXK:1529282.1529414} or will they be processed in a more sophisticated way? That can mean incorporating lexical and syntax analyses, known from creation of compilers, or even a form of an analysis of semantics \cite{compilers}.

The result of lexical and syntax analyses can be some kind of so-called syntax tree \cite{compilers}. It is a structure representing a word according to a formal grammar of some language. In our case, the language is XQuery, its grammar is defined in REF \todo{ref} and every query is a word of the XQuery language. Leaves of the tree represent terminals of the grammar while inner nodes represent non-terminals. \todo{priklad nejakeho stromu?} From the point of view of this work, the syntax tree can be perceived as a preprocessed form of a query keeping its complete meaning and making its further processing more convenient. For instance, the tree can simplify a search for FLWOR statements. It is transitioned and nodes representing FLWORs are found. Then each subtree determined by one of the found nodes represents a FLWOR statement and it can be analyzed further.

\todo[inline]{Q: definovat grammar, tree, language? A: Pokud bude použito dále, tak ano.}

The syntax tree can be also extended by additional information. An example is a static analysis of expression types. Types of literal expressions are defined, functions have return types, path expressions can return nodes, etc. Types of complex expressions can be determined applying the rules defined in REF \todo{ref}. The inferred expression types can be helpful for example in the analysis of built-in types of nodes as discussed in the previous chapter.

@@

While simplier approaches of the query processing such as the pattern finding limit possibilities of the query utilization, a more complex processing of queries provides a better starting point for consecutive analyses and also for further refinements and additions. Therefore, we decided to incorporate query processing using the lexical and syntax analyses.

@@
